On the presentation of orthogonal groups over polynomial rings

### 2019-01-23 dl-paper.tex
Proved injectivity lemma

Schlichting gave a more general definition which still agrees with the classical one in the affine case.
In particular one has:
\begin{equation}
 \GW_i^{[k]}(R) = \begin{array}{cl} \mathrm{KO}_i(R), & k = 0, \\ U_i(R), & k = 1, \\ \mathrm{KSp}_i(R), & k = 2, \\ {}_{-1}\!U_i(R), & k = 3. \end{array}
\end{equation}

\begin{lemma}\label{lem:bass-ft} Suppose that $2 \in R^*$, then for any $i\geq 1$, $k\in \ZZ/4\ZZ$ one has
 \[ \GW_i^{[k]}(R[X, X^{-1}]) \cong \GW_i^{[k]}(R) \oplus \GW_{i-1}^{[k-1]}(R). \] \end{lemma}
\begin{proof} This is a very special case of~\cite[Theorem~9.13]{Sch16} and the remark that follows it. \end{proof}

@@

The following lemma summarizes facts about $\K_2(\Phi, -)$ in the field case.
\begin{lemma} \label{lem:field-case}
 Under certain assumptions the map $\St(\Phi, k[t]) \to \St(\Phi, k[t, t^{-1}])$ is injective.
\end{lemma}
\begin{proof}
 %TODO: Add references on Rehmann and Hurrelbrink...
\end{proof}

\begin{prop}
 Suppose that the canonical map $C(\Phi, B, I) \to C(\Phi, R, I)$ is surjective,
  and that Tulenbaev lifting property holds for $\St(\Phi, R)$. 
 Then the canonical map of Steinberg groups $\St(\Phi, B) \to \St(\Phi, R)$ is injective.
\end{prop}
\begin{proof}
 Consider the following commutative diagram with exact rows: 
\begin{equation*} \xymatrix{
 C(\Phi, B, I) \ar[r]^{j_B} \ar@{->>}[d]_{f} & \St(\Phi, B, I) \ar[r]^{\mu_B} \ar[d]_{g} &
 \St(\Phi, B) \ar[r]^{\pi_B} \ar[d]_{h} & \St(\Phi, k[X]) \ar[d]_{i} \\
 C(\Phi, R, I) \ar[r]^{j_R}         & \St(\Phi, R, I) \ar[r]^{\mu_R} \ar@{-->}[ur]^{t}&
 \St(\Phi, R) \ar[r]^{\pi_R}        & \St(\Phi, k[X, X^{-1}]).
}\end{equation*}
Let $a$ be an element of $\Ker(h)$. Since $i$ is injective by~\cref{lem:field-case}, the element $a$ also lies in $\Ker(\pi_B)$ and hence comes from some $b \in \St(\Phi, B, I)$ via $\mu_B$. Since $g(b) \in \Ker(\mu_R)$ there exists some $c \in C(\Phi, R, I)$ such that $j_R(c) = g(b)$. By our assumption, $f$ is surjective, therefore
 $c = f(d)$ for some $d \in C(\Phi, R, I)$ and we obtain the required assertion:
 \[ 1 = \mu_Bj_B(d) = tgj_B(d) = tj_Rf(d) =t(g(b)) = \mu_B(b) = a. \qedhere \]

@@

\begin{theorem} Suppose $A$ is a local ring of characteristic $\neq 2$, then for $\ell \geq 7$
  the canonical map $\St(\rD_\ell, B) \to \St(\rD_\ell, R)$ is injective.
\end{theorem}
\begin{proof}
 In view of the above proposition and~\cref{lem:lemma32} we are left to verify that the map $f\colon C(\rD_\ell, B, I) \to C(\rD_\ell, R, I)$ is surjective. Consider the following commutative diagram, whose rows are given by \cref{characterization}:
 \begin{equation*}\xymatrix{
  \K_3(\rD_\ell, B) \ar[r] \ar[d] & \K_3(\rD_\ell, k[X]) \ar[d] \ar@{->>}[r] & \ar[d]_{f} C(\rD_\ell, B, I) \\
  \K_3(\rD_\ell, R) \ar[r]        & \K_3(\rD_\ell, k[X, X^{-1}]) \ar@{->>}[r]        & C(\rD_\ell, R, I).}\end{equation*}
 From~\cref{lem:bass-ft,lem:summary} we obtain that $\GW_3^{[0]}(k[X, X^{-1}]) \cong \GW_3^{[0]}(k)$.
 In view of~\cref{lem:Panin-stability} the central vertical map is isomorphic to the canonical map
 $\GW_3^{[0]}(k[X]) \to \GW_3^{[0]}(k[X, X^{-1}])$ and therefore is surjective. The requred surjectivity of $f$ now follows.
\end{proof}

>>>>

Schlichting gave a more general definition which still agrees with the classical one in the affine case.
In particular one has:
\begin{equation}
 \GW_i^{[k]}(R) = \left\{\begin{array}{ll} \mathrm{KO}_i(R), & k = 0, \\ U_i(R), & k = 1, \\ \mathrm{KSp}_i(R), & k = 2, \\ {}_{-1}\!U_i(R), & k = 3. \end{array}\right.
\end{equation}

\begin{lemma}\label{bass-ft} Suppose that $2 \in R^*$, then for any $i\geq 1$, $k\in \ZZ/4\ZZ$ one has
 \[ \GW_i^{[k]}(R[X, X^{-1}]) \cong \GW_i^{[k]}(R) \oplus \GW_{i-1}^{[k-1]}(R). \] \end{lemma}
\begin{proof} This is a very special case of~\cite[Theorem~9.13]{Sch16} and the remark that follows it. \end{proof}

@@

\begin{lemma} Let $\Phi$ be a reduced irreducible root system of type $\neq \rG_2$ and $k$ be arbitrary field.
Then $\K_2(\Phi, k[X, X^{-1}]) = \K_2(\Phi, k) \oplus H(\Phi, k)$, where
\[ H(\Phi, k) = \left\{\begin{array}{ll} k^* & \text{if $\Phi$ is nonsymplectic,} \\ \K_1^\mathrm{MW}(k)& \text{otherwise.}  \end{array}\right. \]  \end{lemma}
\begin{proof} In the nonsymplectic case one can choose a long simple root $\alpha$ of $\Phi$ in such a way that there is a commutative diagram of abelian groups
\[\xymatrixcolsep{5pc}\xymatrix{k^* \ar[r]^-{\{ -, X \}_{\alpha}} \ar[rrd]_{\{-, X\}} & \K_2(\Phi, k[X, X^{-1}]) \ar[r] & \K_2(\Phi, k(X)) \ar[d]^{\cong} \\
                                                                                      &                                 & \K_2^\mathrm{M}(k(X)).} \]
Clearly, the diagonal arrow is injective by Milnor's theorem and the vertical arrow is an isomorphism by Matsumoto's theorem~\cite[Theorem~5.10]{Ma69}.
The required assertion now follow from~\cite[Satz~3]{Hur77} and the fact that the left horizontal arrow is injective.

In the symplectic case the assertion follows from~\cite[Theorem~B]{MR91}, provided one verifies that Morita--Rehmann's group $P(F)$ is isomorphic to $\K_1^\mathrm{MW}(F)$,
 which is an easy exercise. Alternatively, this isomorphism can be obtained as a corollary of~\cref{bass-ft} and~\cite[Lemma~4.1.1]{AF17}. \end{proof}

\begin{corollary} \label{field-injectivity} For $\Phi\neq\rG_2$ the map $\St(\Phi, k[X]) \to \St(\Phi, k[X, X^{-1}])$ is injective. \end{corollary}
\begin{proof} The result follows from consideration of the commutative diagram
\[\xymatrix{ & \K_2(\Phi, k) \ar[dl]_{\cong} \ar@{^{(}->}[dr] & \\
               \K_2(\Phi, k[X]) \ar[rr] &               & \K_2(\Phi, k[X, X^{-1}])} \]
in which the left arrow is an isomorphism by the corollary of~\cite[Satz~1]{Re75} and the right arrow is split injective by the above lemma. \end{proof}

\begin{prop}
 Suppose that the canonical map $C(\Phi, B, I) \to C(\Phi, R, I)$ is surjective,
  and that Tulenbaev lifting property holds for $\St(\Phi, R)$. 
 Then the canonical map of Steinberg groups $\St(\Phi, B) \to \St(\Phi, R)$ is injective.
\end{prop}
\begin{proof}
 Consider the following commutative diagram with exact rows: 
\begin{equation*} \xymatrix{
 C(\Phi, B, I) \ar[r]^{j_B} \ar@{->>}[d]_{f} & \St(\Phi, B, I) \ar[r]^{\mu_B} \ar[d]_{g} &
 \St(\Phi, B) \ar[r]^{\pi_B} \ar[d]_{h} & \St(\Phi, k[X]) \ar[d]_{i} \\
 C(\Phi, R, I) \ar[r]^{j_R}         & \St(\Phi, R, I) \ar[r]^{\mu_R} \ar@{-->}[ur]^{t}&
 \St(\Phi, R) \ar[r]^{\pi_R}        & \St(\Phi, k[X, X^{-1}]).
}\end{equation*}
Let $a$ be an element of $\Ker(h)$. Since $i$ is injective by~\cref{field-injectivity}, the element $a$ also lies in $\Ker(\pi_B)$ and hence comes from some $b \in \St(\Phi, B, I)$ via $\mu_B$. Since $g(b) \in \Ker(\mu_R)$ there exists some $c \in C(\Phi, R, I)$ such that $j_R(c) = g(b)$. By our assumption, $f$ is surjective, therefore
 $c = f(d)$ for some $d \in C(\Phi, R, I)$ and we obtain the required assertion:
 \[ 1 = \mu_Bj_B(d) = tgj_B(d) = tj_Rf(d) =t(g(b)) = \mu_B(b) = a. \qedhere \]

@@

\begin{theorem} Suppose $A$ is a local ring of characteristic $\neq 2$, then for $\ell \geq 7$
  the canonical map $\St(\rD_\ell, B) \to \St(\rD_\ell, R)$ is injective.
\end{theorem}
\begin{proof}
 In view of the above proposition and~\cref{lem:lemma32} we are left to verify that the map $f\colon C(\rD_\ell, B, I) \to C(\rD_\ell, R, I)$ is surjective. Consider the following commutative diagram, whose rows are given by \cref{characterization}:
 \begin{equation*}\xymatrix{
  \K_3(\rD_\ell, B) \ar[r] \ar[d] & \K_3(\rD_\ell, k[X]) \ar[d] \ar@{->>}[r] & \ar[d]_{f} C(\rD_\ell, B, I) \\
  \K_3(\rD_\ell, R) \ar[r]        & \K_3(\rD_\ell, k[X, X^{-1}]) \ar@{->>}[r]        & C(\rD_\ell, R, I).}\end{equation*}
 From~\cref{bass-ft,lem:summary} we obtain that $\GW_3^{[0]}(k[X, X^{-1}]) \cong \GW_3^{[0]}(k)$.
 In view of~\cref{lem:Panin-stability} the central vertical map is isomorphic to the canonical map
 $\GW_3^{[0]}(k[X]) \to \GW_3^{[0]}(k[X, X^{-1}])$ and therefore is surjective. The requred surjectivity of $f$ now follows.
\end{proof}
